---
title: "GEAC Notebook"
author:
- name: Oliver M. Crook
- name: William J. Bradshaw
output:
  BiocStyle::html_document:
    toc_float: yes
---
# Introduction

This is the workbook accompanying the manuscript: Analysis of the first Genetic
Engineering Attribution Challenge. This workbook allows the exploration
of the underlying data and the reproduction of all the figures. Each section
is designed to be self-contained so you can use the table of contents bar
on the left hand side to scroll down to a figure of interest. 


# Figure 1B
#==============================================================================
# MAIN TEXT FIGURE 1B: HISTORICAL GEA ACCURACY
#==============================================================================

## Libraries & formatting
```{r,}
source("scripts/aux_format-plots.R")

# Input paths
path_key <- "data/key-model-stats.csv"

# Output path
out_path  <- "figures/main_fig1-accuracy.svg"

# Define plotting theme
xtext = element_text(angle = 45, vjust = 1, hjust = 1,
                     margin = lmargin(t=0.3,b=-0.9),
                     size = fontsize_title)
ltext = element_text(margin = lmargin(t=0.1,b=0.1,l=0.1,r=0))
theme_col <- theme_internal_strong + theme(
  axis.title.x = element_blank(),
  axis.text.x.bottom = xtext,
  legend.text = ltext,
  legend.position = c(0.98, 0.98),
  legend.justification = c("right", "top"),
  legend.title = element_blank(),
  legend.margin = lmargin(rep(0.1,4)),
  legend.box.margin = lmargin(rep(0.1,4)),
  legend.key.height = lines(0.9),
  legend.key.width  = lines(0.9),
  aspect.ratio = NULL,
  plot.margin = lmargin(t=0.3,l=0.05,r=0.05)
)
```

## Set parameters
```{r,}

out_width <- 2.762
out_asp   <- 2.050/out_width
source_names_raw <- c("BLAST (new)", "BLAST (Alley et al., original)",
                      "BLAST (Alley et al., corrected)", "Nielsen & Voigt (2018)", "Alley et al. (2020)",
                      "GEAC (1st place)", "GEAC (2nd place)", "GEAC (3rd place)",
                      "GEAC (4th place)", "GEAC Ensemble")
source_names_plot <- c(NA, NA, "BLAST", "Nielsen & Voigt\n(2018)",
                       "Alley et al.\n(2020)", "GEAC\n(1st place)", NA, NA, NA,
                       "GEAC\nEnsemble")
source_levels <- source_names_plot[!is.na(source_names_plot)]
palette_fill <- unname(c(palette_primary_dark["grey"], palette_primary["grey"]))
```


```{r,}
#------------------------------------------------------------------------------
# Import and process data
#------------------------------------------------------------------------------

tab_sources <- tibble(source = source_names_raw,
                      source_new = source_names_plot)

data_key <- import_csv(path_key) %>%
  inner_join(tab_sources, by = "source") %>%
  filter(!is.na(source_new)) %>% select(-source) %>%
  select(source = source_new, everything()) %>%
  mutate(source = factor(source, levels = source_levels)) %>%
  arrange(source)

tab_error <- data_key %>% select(source, accuracy_top1, accuracy_top10) %>%
  gather(metric, accuracy, -source) %>%
  mutate(metric = sub("accuracy_top", "Top ", metric),
         metric = fct_inorder(metric),
         error = 1-accuracy)
```

```{r,}
#------------------------------------------------------------------------------
# Make headline plot
#------------------------------------------------------------------------------

g_acc <- ggplot(tab_error, aes(x=source, y=accuracy, fill = metric)) +
  geom_col_mid(position = "dodge") +
  scale_y_percent(name = "Accuracy (%)") +
  scale_fill_manual(values = palette_fill) +
  theme_col + theme(legend.position = c(0.95, 0.05),
                    legend.justification = c("right", "bottom"))

g_error <- ggplot(tab_error, aes(x=source, y=error, fill=metric)) +
  geom_col_mid(position = "dodge") +
  scale_y_cont_nominor(name = "Misclassification rate (%)",
                       breaks = seq(0,1,0.1), limits = c(0,0.58),
                       expand = c(0,0), labels = function(x) x*100) +
  scale_fill_manual(values = palette_fill) +
  theme_col

g_error
```

# Figure 1C
#==============================================================================
# MAIN TEXT FIGURE 1C, PART 1: DATASET SIZES & SEQUENCE LENGTHS
#==============================================================================

```{r,}
# Libraries
source("scripts/aux_format-plots.R")

# Data paths
size_path <- "data/dataset-size.csv"
len_path  <- "data/sequence-lengths-sample.csv"

# Output path
out_svg <- "figures/main_fig1-datasets.svg"

# Plotting parameters
asp_ratio  <- 0.6 # Aspect ratio for export (width/height)

# Scales
scale_fill_palette <- purrr::partial(scale_fill_manual,
                                     values = unname(palette_primary))

# Factor levels
dataset_levels <- c("train", "leaderboard", "test")
```

```{r,}
#------------------------------------------------------------------------------
# Import and process data
#------------------------------------------------------------------------------

# Dataset sizes
data_size <- import_csv(size_path) %>% 
  mutate(dataset = factor(dataset, levels = dataset_levels))

# Sampled sequence lengths from each dataset
data_len <- import_csv(len_path) %>% 
  mutate(dataset = factor(dataset, levels = dataset_levels))
```

```{r,}
#------------------------------------------------------------------------------
# Plot relative sizes (for metadata / labels columns)
#------------------------------------------------------------------------------

g_size <- ggplot(data_size, aes(x=1, y=size, fill = dataset)) +
  geom_col(position = "fill") + scale_fill_palette() +
  theme_void() + theme(legend.position = "blank")
```

```{r,}
#------------------------------------------------------------------------------
# Plot sequence lengths (for sequence data column)
#------------------------------------------------------------------------------

g_seq <- ggplot(data_len, 
                aes(x=sequence_order, y=sequence_length, fill=dataset)) +
  geom_col(width = 1) + 
  coord_flip() + 
  scale_y_reverse() + scale_x_reverse() +
  scale_fill_palette() +
  theme_void() + theme(legend.position = "blank")
```

```{r,}
#------------------------------------------------------------------------------
# Arrange & save
#------------------------------------------------------------------------------

# Arrange with patchwork
g_all <- g_seq + g_size + g_size + plot_layout(widths = c(12,1,1))
g_all

#*Note that this figure is eddited externally to produce the figure in the 
#*manuscript.

```

# Figure 2
#==============================================================================
# MAIN TEXT FIGURE 2: CORE COMPETITION METRICS
#==============================================================================

## Libraries & formatting
```{r,}
source("scripts/aux_format-plots.R")

# Input path
path_countries <- "data/country-percentages.csv"
path_teams     <- "data/team-stats.csv"
path_accuracy  <- "data/team-top-n-accuracy.csv"

# Output paths
out_path_main <- "figures/main_fig2.png" # Main figure
out_path_si_nsub <- "figures/si_n-submissions.png" # SI figure on submission #s
out_path_si_acc_all <- "figures/si_accuracy-all-submissions.png"
out_path_si_f1_all <- "figures/si_f1-all-submissions.png"
out_path_si_pr_decile <- "figures/si_precision-recall-deciles.png"

# Output parameters
output_width_in <- page_width_in # Width of page
output_width_small <- fig_width_single_wide_in # Width for wide single-panel SI figures
output_asp_main <- 0.7 # Main-text figure
output_asp_si_nsub <- 0.55 # Two-panel n_submissions figure
output_asp_si_acc <- 0.7 # One-panel accuracy figure
output_asp_si_f1 <- 0.7 # One-panel F1 figure
output_asp_si_pr_decile <- 0.6
```

```{r,}
#------------------------------------------------------------------------------
# Define palettes, scales, themes
#------------------------------------------------------------------------------

# Palettes
palette_countries_part <- c(rbind(palette_primary_dark[c("blue", "orange", "green2")],
                                  palette_primary_pale[c("yellow", "pink", "blue")]),
                            palette_primary_dark["beige"],palette_primary_mid["grey"])

#------------------------------------------------------------------------------
# Import data
#------------------------------------------------------------------------------

# Country data (for pie chart)
data_countries <- suppressMessages(read_csv(path_countries))

# Team data
data_teams <- suppressMessages(read_csv(path_teams))

# Accuracy data
data_accuracy <- as.matrix(read.csv(path_accuracy, header = FALSE))


```

## Designate country display groups
```{r,}
#------------------------------------------------------------------------------
# Make pie chart
#----------------------------`--------------------------------------------------
europe_countries <- c("UK", "Netherlands", "Germany", "France", "Spain",
                      "Other Europe")
asia_countries   <- c("China", "Other Asia")
other_countries <- c("Other", "Other North America", "Other South America",
                     "Africa", "Oceania")
groups_out <- c("USA", "India", "Russia", "Other\nEurope", "Other\nAsia",
                "Canada", "Brazil", "Other")
data_countries_collapsed <- data_countries %>%
  mutate(in_europe = country %in% europe_countries,
         in_asia   = country %in% asia_countries,
         in_other  = country %in% other_countries,
         group = ifelse(in_europe, "Other\nEurope",
                        ifelse(in_asia, "Other\nAsia",
                               ifelse(in_other, "Other", country)))) %>%
  group_by(group) %>%
  summarise(percent_participants = sum(percent_participants),
            percent_site_visitors = sum(percent_site_visitors)) %>%
  rename(country = group) %>%
  arrange(match(country, groups_out)) %>%
  mutate(country = fct_inorder(country))

names(palette_countries_part) <- groups_out

# Registered participant data
data_countries_part <- data_countries_collapsed %>%
  rename(percent = percent_participants) %>%
  mutate(label = country,
         csum = rev(cumsum(rev(percent))),
         pos = percent/2 + lead(csum, 1),
         pos = if_else(is.na(pos), percent/2, pos)) %>%
  select(-percent_site_visitors, -csum)

# Text alignment
hjust_countries_part <- c(0, 0, 0, 1.1, 1, 1, 0.7, 0.2)
vjust_countries_part <- c(0, 0.5, 1, 0.7, 0.8, 0.6, 0, 0)

# Make pie charts
g_countries_part <- ggplot(data_countries_part,
                           aes(x=1, y=percent, fill=fct_inorder(label))) +
  geom_col(width = 1) +
  coord_polar(theta = "y", direction = -1, clip = "off") +
  geom_text(aes(y = pos, label = label), x=1.8,
            size = rescale_font(fontsize_base),
            hjust = hjust_countries_part,
            vjust = vjust_countries_part, lineheight = 0.9) +
  geom_segment(aes(y=pos, yend=pos), x=1.5, xend=1.7,
               size = 0.2) +
  scale_y_continuous(name = "Participant countries",
                     breaks = data_countries_part$pos,
                     labels = data_countries_part$label) +
  scale_fill_manual(values = palette_countries_part) +
  guides(fill = FALSE) +
  theme_blank + theme(
    axis.title = NULL,
    axis.title.x = element_text(margin = lmargin(1.5,0,0,0),
                                face = "bold", hjust = 0.5,
                                vjust = 0.5),
    axis.title.y = element_blank(),
    aspect.ratio = 1,
    plot.margin = lmargin(r=2.5)
  )
```


```{r,}
#------------------------------------------------------------------------------
# Rank vs number of submissions (main plots)
#------------------------------------------------------------------------------

# Prepare dataset
data_prf <- data_teams %>% select(-starts_with("accuracy")) %>%
  mutate(ranking_pc = prediction_track_ranking/max(prediction_track_ranking),
         ranking_decile = ceiling(ranking_pc*10)) %>%
  mutate(winning_rank = ifelse(prediction_track_ranking <= n_winners,
                               prediction_track_ranking, "Other"),
         winning_rank = factor(winning_rank, levels = c(1:n_winners, "Other")),
         )
data_prf_winners    <- data_prf %>% filter(winning_rank != "Other")
data_prf_nonwinners <- data_prf %>% filter(winning_rank == "Other")

# Make complete plot
g_nsub_all <- ggplot(aes(x=prediction_track_ranking, y=n_entries,
                         colour=winning_rank), data = NULL) +
  geom_line_thin(colour = last(palette_winners), data = data_prf) +
  geom_point_nonwinners(data = data_prf_nonwinners) +
  geom_point_winners(data = data_prf_winners) +
  scale_x_ranking_all_trunc() +
  scale_y_cont_nominor(name = "# Submissions", limits = c(0, 150),
                       expand = c(0,0)) +
  scale_colour_winners() +
  theme_base

# Top 100 only
g_nsub_100 <- g_nsub_all + scale_x_ranking_100()
```


```{r,}
#------------------------------------------------------------------------------
# Rank vs number of submissions (decile plot)
#------------------------------------------------------------------------------

# Derive decile labels
data_prf_deciles_labelled <- data_prf %>%
  group_by(ranking_decile) %>%
  arrange(prediction_track_ranking) %>%
  mutate(decile_rank_min = min(prediction_track_ranking),
         decile_rank_max = max(prediction_track_ranking),
         decile_label = paste(decile_rank_min, 
                              decile_rank_max, sep=" to "),
         decile_label_long = paste("Rank", decile_label),
         decile_label = fct_inorder(decile_label),
         decile_label_long = fct_inorder(decile_label_long))

# Summarise by decile
data_prf_deciles_summ <- data_prf_deciles_labelled %>%
  summarise(avg_n_entries = mean(n_entries),
            decile_label = decile_label[1])

# Make decile plot
g_nsub_deciles <- ggplot(data_prf_deciles_summ, 
                         aes(x=decile_label, y=avg_n_entries)) +
  geom_col_mid(fill = last(palette_winners)) +
  scale_x_discrete(name = "Prediction Track ranking") +
  scale_y_cont_nominor(name = "Mean submissions per team",
                       expand = c(0,0), limits = c(0,50)) +
  theme_col
```

```{r,}
#------------------------------------------------------------------------------
# Accuracy vs rank
#------------------------------------------------------------------------------

# Prepare data for plotting
tab_rank_accuracy <- data_teams %>%
  select(prediction_track_ranking, starts_with("accuracy_")) %>%
  gather(type, accuracy, -prediction_track_ranking) %>%
  mutate(N = as.integer(sub("accuracy_top", "", type)),
         type = paste("Top", N),
         type = factor(type, levels = paste("Top", sort(unique(N)))),
         error = 1 - accuracy) %>%
  arrange(type == "Top 1", type == "Top 10", type == "Top 5")

# Make plot (all teams)
g_accuracy_rank_all <-  ggplot(tab_rank_accuracy, 
                               aes(x=prediction_track_ranking, y=accuracy,
                                   colour=type, shape=type)) +
  geom_line_thin() + geom_point(size = 1) +
  scale_x_ranking_all() + scale_y_percent(name = "Accuracy (%)") +
  scale_colour_accuracy() + scale_shape_multi(name = "Accuracy metric:") +
  theme_base + theme(aspect.ratio = 1/2)

# Top 100 only
g_accuracy_rank_100 <- g_accuracy_rank_all + scale_x_ranking_100() +
  theme(aspect.ratio = plot_aspect_ratio_base)
```


```{r,}
#------------------------------------------------------------------------------
# Macro precision, recall, F1 (all submissions)
#------------------------------------------------------------------------------

# Split labelled dataset by winner status (need for facet plot later)
data_prf_dl_winners    <- data_prf_deciles_labelled %>%
  filter(winning_rank != "Other")
data_prf_dl_nonwinners <- data_prf_deciles_labelled %>%
  filter(winning_rank == "Other")

# Make precision/recall plot
g_precision_recall <- ggplot(aes(x=macro_precision, y=macro_recall,
                                 colour=winning_rank), data = NULL) +
  geom_abline(size = 0.3, colour = palette_primary_mid["grey"],
              linetype = "dashed") +
  geom_point_nonwinners(data = data_prf_dl_nonwinners, alpha = 0.7) +
  geom_point_winners(data = data_prf_dl_winners, alpha = 1) +
  scale_x_proportion(name = "Precision") +
  scale_y_proportion(name = "Recall") +
  scale_colour_winners() +
  guides(alpha = FALSE) +
  theme_base

# Make macro-F1 plots
g_f1_all <- ggplot(aes(x=prediction_track_ranking, y=macro_f1,
                       colour=winning_rank), data = NULL) +
  geom_line_thin(data = data_prf, colour = last(palette_winners)) +
  geom_point_nonwinners(data = data_prf_nonwinners) +
  geom_point_winners(data = data_prf_winners) +
  scale_x_ranking_all() + scale_y_proportion(name = "F1 score") +
  scale_colour_winners() +
  theme_base + theme(aspect.ratio = 1/2)

g_f1_100 <- g_f1_all + scale_x_ranking_100() +
  theme(aspect.ratio = plot_aspect_ratio_base)
```

```{r,}
#------------------------------------------------------------------------------
# Precision/recall by decile
#------------------------------------------------------------------------------

# Make plot
g_precision_recall_deciles <- g_precision_recall +
  facet_wrap(~decile_label_long, ncol = 5) +
  theme_col + theme(aspect.ratio = 1,
                    strip.text = element_text(face = "plain"))
```

```{r,}
#------------------------------------------------------------------------------
# Winning team accuracy curves
#------------------------------------------------------------------------------

# Sort accuracy scores by top-10 accuracy, then filter for winners
acc_order <- order(data_accuracy[,10], decreasing = TRUE)
n_winners <- 4
n_labs <- ncol(data_accuracy)
data_accuracy_ordered <- data_accuracy[acc_order,]
data_accuracy_winners <- data_accuracy_ordered[1:n_winners,]
dimnames(data_accuracy_winners) <- list(rank = 1:n_winners, N = 1:n_labs)

# Convert into tibble
tab_accuracy_winners <- data_accuracy_winners %>% 
  as.table(responseName = "accuracy", stringsAsFactors = FALSE) %>% 
  as.data.frame(stringsAsFactors = FALSE, responseName = "accuracy") %>% 
  as_tibble %>% mutate(N = as.numeric(N))

# Plot accuracy
g_accuracy_winners <- ggplot(tab_accuracy_winners, 
                             aes(x=N, y=accuracy, colour = rank)) +
  geom_line_thin() + geom_point_winners() +
  scale_x_log10(name = "N") +
  scale_y_cont_nominor(name = "Top-N accuracy (%)",
                     labels = function(x) as.integer(x*100),
                     breaks = seq(0,1,0.1), limits = c(0.6, 1),
                     expand = c(0,0.01)) +
  scale_colour_winners() +
  theme_base
```

## Bring all figures together
```{r,}
#------------------------------------------------------------------------------
# Collect all subfigures for main figure
#------------------------------------------------------------------------------

# Collate subfigures
g_acc_win_out <- g_accuracy_winners + guides(colour = FALSE)
g_pre_rec_out <- g_precision_recall + guides(colour = FALSE, alpha = FALSE)

# Assemble grid
g_figure <- (g_countries_part + g_nsub_100 + g_accuracy_rank_100) / 
  (g_acc_win_out + g_pre_rec_out + g_f1_100) +
  plot_annotation(tag_levels = "a") +
  plot_layout(guides = "collect", #ncol = 3, nrow = 2, 
              widths = c(1,1,1), heights = c(1,1)
              ) & 
  theme_legend & theme(
    legend.box = "vertical"
    )
g_figure

```

```{r,}
#------------------------------------------------------------------------------
# Collate SI figures
#------------------------------------------------------------------------------

g_si_nsub <- wrap_plots(g_nsub_all, g_nsub_deciles, ncol = 2, widths = c(1,1),
                        guides = "collect") +
  plot_annotation(tag_levels = "a") &
  theme_legend
g_si_nsub
```

# Figure 3

#==============================================================================
# MAIN TEXT FIGURE 3 & 4: X-metrics
#==============================================================================

```{r,}
# Libraries & formatting
source("scripts/aux_format-plots.R")

# Input paths
path_accuracy <- "data/team-top-n-accuracy.csv"
path_teams <- "data/team-stats.csv"
path_key <- "data/key-model-stats.csv"

# Output paths
out_path_fig3_main <- "figures/main_fig3.png"
out_path_fig4_main <- "figures/main_fig4.png"
out_path_x_all <- "figures/si_xmetrics-all-submissions.png"

# Parameters
example_team_rank <- 126 # Ranking of team to use for example plot
tag_margin_fig3 <- lmargin(t = 0.2, r = 0.2, b = -0.5, l = 0.3) # Margin of subfigure labels
tag_margin_fig4 <- lmargin(t = 0, r = 0.2, b = 0.5, l = 0.3)
source_names_raw <- c("BLAST (new)", "BLAST (Alley et al., original)",
                      "BLAST (Alley et al., corrected)", "Nielsen & Voigt (2018)", "Alley et al. (2020)",
                      "GEAC (1st place)", "GEAC (2nd place)", "GEAC (3rd place)",
                      "GEAC (4th place)", "GEAC Ensemble")
source_names_plot <- c(NA, NA, "BLAST", "Nielsen & Voigt (2018)",
                       "Alley et al. (2020)", "GEAC (1st place)", NA, NA, NA,
                       "GEAC Ensemble")

# Output parameters
out_width_main <- page_width_in
out_width_si   <- fig_width_single_wide_in
out_asp_main   <- 0.45
out_asp_si     <- 0.7

# Palettes
palette_nlabs <- unname(palette_primary_darkmid["grey"])
palette_example <- unname(palette_primary["green2"])

# Factor levels
x_levels <- c("X99", "X95", "X90", "X80")
win_levels <- c(1:n_winners, "Other")
outlier_levels <- c(TRUE, FALSE) # Fig. 4 outliers
source_levels <- source_names_plot[!is.na(source_names_plot)]

# Scales
scale_shape_x <- purrr::partial(scale_shape_multi,
                                name = "Metric:")
```

```{r,}
#------------------------------------------------------------------------------
# Read data
#------------------------------------------------------------------------------

# Import accuracy and team stats
data_accuracy <- import_csv_matrix(path_accuracy)
data_teams    <- import_csv(path_teams)

# Import, rename and filter key sources
tab_sources <- tibble(source = source_names_raw,
                      source_new = source_names_plot)
data_key <- import_csv(path_key) %>%
    inner_join(tab_sources, by = "source") %>%
    filter(!is.na(source_new)) %>% select(-source) %>%
    select(source = source_new, everything()) %>%
    mutate(source = factor(source, levels = source_levels)) %>%
    arrange(source)
```

```{r,}
#-----------------------------------------------------------------------------
# X95 overview plot (example team)
#------------------------------------------------------------------------------

# Get example data
tab_example <- filter(data_teams, 
                      prediction_track_ranking == example_team_rank)
which_accuracy <- which(data_accuracy[,10] == tab_example$accuracy_top10 &
                            data_accuracy[,5] == tab_example$accuracy_top5)
accuracies_example <- tibble(N = 1:ncol(data_accuracy),
                             accuracy = data_accuracy[which_accuracy,])

# Extract annotation data
annot_scale <- 1.15
tab_annot <- tibble(
    y = c(0.95, 0.99),
    x = c(tab_example$X95, tab_example$X99),
    colour = palette_x[c(2,1)],
    label = c("X95", "X99"),
    hjust = c(1,0)
) %>% mutate(x_scaled = x * c(1/annot_scale, annot_scale))

# Plot helper functions & parameters
geom_seg <- purrr::partial(geom_segment, size = 0.3, linetype = "dashed",
                           data = tab_annot)

#ggplot
g_example <- ggplot(accuracies_example, aes(x = N, y = accuracy))  +
    labs(x = "N", y = "Top-N Accuracy (%)", title = "") +
    geom_line_thin(color = palette_example[1]) +
    geom_point_fixed(color = palette_example[1], size = 1) +
    geom_seg(aes(xend = x, y = y, yend = y, colour = colour), x = -Inf) +
    geom_seg(aes(yend = y, x = x, xend = x, colour = colour), y = -Inf) +
    geom_text(aes(x = x_scaled, colour = colour, hjust = hjust, 
                  label = label), data = tab_annot, y = 0.51,
              size = rescale_font(fontsize_base), vjust = 0) +
    scale_y_cont_nominor(labels = function(x) as.integer(x*100),
                         breaks = seq(0,1,0.1), limits = c(0.5, 1),
                         expand = c(0,0.01)) +
    scale_x_log10(limits = c(NA,2500)) +
    scale_colour_identity() +
    theme_base + theme(legend.position = "none",
                       plot.tag = element_text(margin = tag_margin_fig3))

g_example
```

```{r,}
#-----------------------------------------------------------------------------
# XN scores plot (main figure)
#------------------------------------------------------------------------------

# Prepare dataset
tab_exclusion <- data_teams %>% rename(ranking = prediction_track_ranking) %>%
    select(ranking, starts_with("X")) %>%
    gather(metric, value, -ranking) %>%
    mutate(metric = factor(metric, levels = x_levels))

# Make plot
g_exclusion_100 <- ggplot(filter(tab_exclusion, metric %in% c("X99", "X95")),
                          aes(x = ranking, y = value, color = metric,
                              shape = metric)) +
    geom_point(size = 1) +
    geom_hline(yintercept = n_labs, colour = palette_nlabs[1],
               linetype = "dashed", size = 0.3) +
    scale_x_ranking_100() +
    scale_y_log10(name = "N", minor_breaks = NULL, limits = c(1, n_labs*1.1),
                  expand = c(0,0)) +
    scale_colour_x() + scale_shape_x() +
    theme_base + theme(plot.tag = element_text(margin = tag_margin_fig3))

g_exclusion_100

#-----------------------------------------------------------------------------
# XN scores plot (SI)
#------------------------------------------------------------------------------

# Make plot
g_exclusion_all <- ggplot(tab_exclusion,
                          aes(x = ranking, y = value, color = metric,
                              shape = metric)) +
    geom_point(size = 1) +
    geom_hline(yintercept = n_labs, colour = palette_nlabs[1],
               linetype = "dashed", size = 0.3) +
    scale_x_ranking_all() +
    scale_y_log10(name = "N", limits = c(1, NA),
                  expand = c(0,0)) +
    scale_colour_x() + scale_shape_x() +
    theme_base + theme(plot.tag = element_text(margin = tag_margin_fig3),
                       aspect.ratio = 1/2)

g_exclusion_all

#------------------------------------------------------------------------------
# Key model XN (summary, full in SI from another script)
#------------------------------------------------------------------------------

df <- data_key %>% select(source, X99, X95) %>% gather(metric, N, -source) %>%
    mutate(metric = factor(metric, levels = x_levels))

df_annot <- data_key %>% select(source, N=X99) %>%
    filter(grepl("GEAC", source)) %>% 
    mutate(metric="X99")

g_prior <- ggplot(df, aes(x = source, y = N, fill = metric)) +
    geom_col_mid(position = "dodge") +
    geom_hline(yintercept = n_labs, colour = palette_nlabs[1],
               linetype = "dashed", size = 0.3) +
    annotate("text", x = Inf, y = n_labs-40, label = "All labs",
             size = rescale_font(fontsize_base), hjust = 1, vjust = 1,
             colour = palette_nlabs[1]) +
    geom_text(aes(colour = metric, label = N), data = df_annot,
              size = fontsize_base_rescaled, angle = 45,
              hjust = 1, vjust = 1, nudge_y = 320, nudge_x = 0.05,
              show.legend = FALSE) +
    scale_y_cont_nominor(name = "N", limits = c(0,1400),
                         breaks = seq(0,2000,400), expand = c(0,0)) +
    scale_fill_x() + scale_colour_x_dark() +
    theme_col + theme(axis.title.x.bottom = element_blank(),
                      plot.tag = element_text(margin = tag_margin_fig3))

g_prior
```

# Figure 4
```{r,}
#------------------------------------------------------------------------------
# Linear exclusion plots (Fig. 4)
#------------------------------------------------------------------------------

# Define outlier values
outlier_min_100 <- list(X95 = n_labs-20, X99 = n_labs-20)

# Add winner annotations to exclusion data
tab_exclusion_linear <- tab_exclusion %>% 
    filter(metric %in% c("X99", "X95")) %>%
    mutate(winning_rank = ifelse(ranking <= n_winners, ranking, "Other"),
           winning_rank = factor(winning_rank, levels = win_levels))

# Split by metric
tab_el_split <- tab_exclusion_linear %>%
    group_by(metric) %>% group_split() %>%
    setNames(sapply(., function(x) first(x$metric))) %>%
    lapply(., function(x) select(x, -metric))

# Filter to top-100
tab_el_split_100 <- lapply(tab_el_split, function(x) filter(x, ranking <= 100))

# Assign outlier status
tab_els_outliers_100 <- lapply(names(tab_el_split_100), function(x)
    tab_el_split_100[[x]] %>% 
        mutate(outlier = value >= outlier_min_100[[x]],
               outlier = factor(outlier, levels = outlier_levels))) %>%
    setNames(names(tab_el_split_100))

# Define basic plotting function (without axis scales)
plot_exc_linear <- function(data, metric, outlier_pad = 5,
                            force_values = 0){
    # Split data by winner status
    data_winners <- data %>% filter(winning_rank != "Other")
    data_nonwinners <- data %>% filter(winning_rank == "Other")
    # Generate outlier padding data (to prevent zero-height facets)
    # Make plot
    a <- aes(x=ranking, y=value, colour=winning_rank)
    g <- ggplot(data = NULL, mapping = a) +
        geom_point_nonwinners(data = data_nonwinners) +
        geom_point_winners(data = data_winners) +
        facet_grid(outlier~., scales = "free_y", space = "free_y") +
        scale_colour_winners() + labs(y = metric) +
        theme_base + 
        theme(strip.text = element_blank(), aspect.ratio = NULL,
              plot.tag = element_text(margin = tag_margin_fig4))
    # Force particular values if needed
    if (length(force_values) > 0){
        data_force <- tibble(ranking = n_winners + 1, value = force_values,
                             outlier = factor(FALSE, levels = outlier_levels),
                             winning_rank = "Other")
        g <- g + geom_point_nonwinners(data = data_force, alpha = 0)
    }
    # Add outlier padding if needed
    if (sum(data$outlier == TRUE) > 0){
        data_outliers <- data %>% filter(outlier == TRUE)
        pad_values <- c(min(data_outliers$value)-outlier_pad,
                        max(data_outliers$value)+outlier_pad)
        data_outlier_pad <- tibble(ranking = n_winners + 1, value = pad_values,
                                   outlier = factor(TRUE, levels = outlier_levels),
                                   winning_rank = "Other")
        g <- g + geom_point_nonwinners(data = data_outlier_pad, alpha = 0)
    }
    return(g)
}

# Make top-100 plots
outlier_pad_100 <- list(X95 = 50/3, X99 = 50)
y_breaks_100 <- list(X95 = c(seq(0, n_labs-100, 100), n_labs),
                     X99 = c(seq(0, n_labs, 300), n_labs))
force_values_100 <- list(X95 = c(0,400), X99 = c(0,1200))
g_exc_linear_100 <- lapply(names(tab_els_outliers_100), function (x)
    plot_exc_linear(tab_els_outliers_100[[x]], x, outlier_pad_100[[x]],
                    force_values_100[[x]]) +
        scale_x_ranking_100() + 
        scale_y_cont_nominor(breaks = y_breaks_100[[x]],
                             expand = c(0,0))) %>%
    setNames(names(tab_els_outliers_100))

#-----------------------------------------------------------------------------
# Bring plots together
#------------------------------------------------------------------------------

# Fig. 3
fig3 <- wrap_plots(g_example, g_exclusion_100, g_prior + guides(fill = FALSE),
                   ncol = 3) +
    plot_annotation(tag_levels = 'a') + plot_layout(guides = "collect") &
    theme_legend & theme(legend.box.spacing = lines(-0.3),
                         plot.margin = lmargin(c(0, 0.1, 0.1, 0)))

fig3

# Fig. 4 main
fig4 <- wrap_plots(g_exc_linear_100[["X99"]], g_exc_linear_100[["X95"]],
                   ncol = 2) +
    plot_annotation(tag_levels = 'a') + plot_layout(guides = "collect") &
    theme_legend & theme(plot.margin = lmargin(c(0.2, 0.2, 0.1, 0)))

fig4
```

# Supplementary Figure: Calibration

#==============================================================================
# SI FIGURE: CALIBRATION
#==============================================================================

```{r,}
# Libraries & formatting
source("scripts/aux_format-plots.R")

# Input path
path_teams <- "data/team-stats.csv"

# Output parameters
out_path <- "figures/si_calibration.png"
out_width <- fig_width_single_wide_in
out_asp <- 1.25
out_path_decile <- "figures/si_calibration-deciles.png"
out_asp_decile <- 0.8

# Scales for decile plot
palette_win <- unname(palette_primary_dark[c("green2", "grey")])
palette_win_dec <- c(palette_win[1], rep("black", 10))
scale_fill_win <- purrr::partial(scale_fill_manual, values = palette_win,
                                 name = "Prizewinners?", labels = c("Yes", "No"))
scale_alpha_decile <- purrr::partial(scale_alpha_discrete, range = c(0.45,1),
                              name = "Metric:")
```

```{r,}
#------------------------------------------------------------------------------
# Import and prepare data
#------------------------------------------------------------------------------

data_teams <- import_csv(path_teams) %>% 
    rename(ranking=prediction_track_ranking) %>%
    mutate(winning_rank = ifelse(ranking <= n_winners, ranking, "Other"),
           winning_rank = fct_inorder(winning_rank)) %>%
    select(ranking, winning_rank, exp_calibration_error, max_calibration_error,
           ranking_decile)
data_win <- filter(data_teams, winning_rank != "Other")
data_nonwin <- filter(data_teams, winning_rank == "Other")

#------------------------------------------------------------------------------
# Per-team plots
#------------------------------------------------------------------------------

aes_cal <- purrr::partial(aes, x = ranking, colour = winning_rank)
stat_cspear <- purrr::partial(stat_spear, label.y = 0.98, vjust = 1, hjust = 0,
                              data = data_teams, label.x = 90)

g_mce <- ggplot(aes_cal(y=max_calibration_error), data = NULL) +
    geom_point_winners(data = data_win) +
    geom_point_nonwinners(data = data_nonwin) +
    stat_cspear() +
    scale_colour_winners() + scale_x_ranking_all() +
    scale_y_percent(name = "Maximum Calibration Error (%)") +
    theme_base + theme(aspect.ratio = 1/2)

g_ece <- ggplot(aes_cal(y=exp_calibration_error), data = NULL) +
    geom_point_winners(data = data_win) +
    geom_point_nonwinners(data = data_nonwin) +
    stat_cspear() +
    scale_colour_winners() + scale_x_ranking_all() +
    scale_y_percent(name = "Expected Calibration Error (%)") +
    theme_base + theme(aspect.ratio = 1/2)

g_out <- wrap_plots(g_ece, g_mce, ncol = 1) +
    plot_annotation(tag_levels = "a") + plot_layout(guides = "collect") &
    theme_legend

g_mce
g_ece
g_out

```

```{r,}
#------------------------------------------------------------------------------
# Decile plot
#------------------------------------------------------------------------------

# Define data
data_deciles <- data_teams %>% group_by(ranking_decile) %>%
    summarise(rank_max = max(ranking), rank_min = min(ranking),
              ECE = mean(exp_calibration_error),
              MCE = mean(max_calibration_error)) %>%
    mutate(decile_label = paste(rank_min, rank_max, sep=" to "))
data_winners <- data_teams %>% filter(ranking <= n_winners) %>%
    summarise(ranking_decile = 0,
              decile_label = paste(1, n_winners, sep = " to "),
              rank_min = 1, rank_max = n_winners,
              ECE = mean(exp_calibration_error),
              MCE = mean(max_calibration_error))
data_deciles_plot <- bind_rows(data_winners, data_deciles) %>%
    mutate(decile_label = fct_inorder(decile_label),
           winner = ranking_decile == 0) %>%
    select(-rank_min, -rank_max) %>%
    gather(metric, value, -ranking_decile, -decile_label, -winner) %>%
    arrange(metric) %>% 
    mutate(`Metric:` = fct_inorder(metric),
           winner = factor(winner, levels = c(TRUE, FALSE)))

# Prepare plotting auxiliaries
aes_dec <- aes(x=decile_label, y=value, fill=winner, alpha=metric)

g_deciles <- ggplot(data_deciles_plot, aes_dec) +
    geom_col_mid(position = "dodge") +
    scale_x_discrete(name = "Innovation Track ranking") +
    scale_y_cont_nominor(name = "Mean prediction error (%)",
                         breaks = seq(0,1,0.2), expand = c(0,0),
                         labels = function(x) x*100,
                         limits = c(0,0.8)) +
    scale_fill_win() + scale_alpha_decile() +
    theme_col + theme(
        axis.text.x.bottom = element_text(colour = palette_win_dec),
        aspect.ratio = 1/2
    )
g_deciles
```



# Supplementary Figure: Accuracy and X-Metrics

#==============================================================================
# SI FIGURE: CORRELATIONS BETWEEN ACCURACY AND X METRICS
#==============================================================================

```{r,}
# Libraries & formatting
source("scripts/aux_format-plots.R")

# Input path
path_teams    <- "data/team-stats.csv"

# Output paths
out_path_acc <- "figures/si_accuracy-correlations.png"
out_path_x   <- "figures/si_xmetric-correlations.png"
out_path_mix_all <- "figures/si_mix-correlations-all.png"
out_path_mix_filtered <- "figures/si_mix-correlations-filtered.png"

# Output parameters
out_width <- page_width_in
out_asp   <- 1.05

# Themes & scales
theme_cor <- theme_base + theme(
  axis.title = element_blank(),
  strip.text = element_text(margin = lmargin(rep(0.2, 4))),
  aspect.ratio = 1,
  plot.margin = lmargin(0.3, 0.4, 0.4, 0.4),
)
scale_x_acc <- purrr::partial(scale_x_continuous, breaks = seq(0,1,0.2),
                              minor_breaks = NULL,
                              labels = function(x) round(x*100))
scale_y_acc <- purrr::partial(scale_y_continuous, breaks = seq(0,1,0.2),
                              minor_breaks = NULL,
                              labels = function(x) round(x*100))
scale_x_nlabs <- purrr::partial(scale_x_cont_nominor, limits = c(0,n_labs),
                                breaks = seq(0,2000,400),
                                expand = c(0,0))
scale_y_nlabs <- purrr::partial(scale_y_cont_nominor, limits = c(0,n_labs),
                                breaks = seq(0,2000,400),
                                expand = c(0,0))
```

```{r,}
#------------------------------------------------------------------------------
# Import data
#------------------------------------------------------------------------------

# Team data
data_teams <- suppressMessages(read_csv(path_teams))

#------------------------------------------------------------------------------
# Part 1: Intra-competition accuracy correlations
#------------------------------------------------------------------------------

# Types of accuracy metric
tab_types <- tibble(accuracy_type_base = data_teams %>% 
                      select(starts_with("accuracy")) %>% colnames) %>%
  mutate(accuracy_type_nice = paste(sub("accuracy_top", "Top-", 
                                        accuracy_type_base), "\nAccuracy (%)"),
         accuracy_type_nice_asc = fct_inorder(accuracy_type_nice),
         accuracy_type_nice_desc = fct_rev(accuracy_type_nice_asc)) %>%
  select(-accuracy_type_nice)

# Combinations of metrics to compare
tab_acc_base <- expand_grid(ranking = data_teams$prediction_track_ranking,
                            accuracy_type_1 = tab_types$accuracy_type_base,
                            accuracy_type_2 = tab_types$accuracy_type_base) %>%
  mutate(winning_rank = ifelse(ranking <= n_winners, ranking, "Other"))

# Add nice accuracy names
tab_acc_nice <- tab_acc_base %>%
  inner_join(tab_types,
             by = c("accuracy_type_1" = "accuracy_type_base")) %>% 
  rename(accuracy_nice_1 = accuracy_type_nice_desc) %>%
  select(-accuracy_type_nice_asc) %>%
  inner_join(tab_types,
             by = c("accuracy_type_2" = "accuracy_type_base")) %>% 
  rename(accuracy_nice_2 = accuracy_type_nice_asc) %>%
  select(-accuracy_type_nice_desc)
  
# Add accuracy values
tab_teams_melt <- data_teams %>% 
  select(ranking = prediction_track_ranking, starts_with("accuracy")) %>%
  gather(accuracy_type, accuracy, -ranking)
tab_acc <- tab_acc_nice %>%
  inner_join(tab_teams_melt,
             by = c("ranking" = "ranking", 
                    "accuracy_type_1" = "accuracy_type")) %>% 
  rename(accuracy_1 = accuracy) %>%
  inner_join(tab_teams_melt,
             by = c("ranking" = "ranking",
                    "accuracy_type_2" = "accuracy_type")) %>% 
  rename(accuracy_2 = accuracy)

# Split into winners/nonwinners
tab_acc_win <- tab_acc %>% filter(winning_rank != "Other")
tab_acc_nonwin <- tab_acc %>% filter(winning_rank == "Other")

# Make plots
aes_col <- aes(colour=winning_rank)
stat_spear_no_p <- purrr::partial(stat_spear, digits = 3,
                                  mapping = aes(label = ..r.label..))
stat_aspear <- purrr::partial(stat_spear_no_p, label.x = 0.05, label.y = 0.95,
                              vjust = 1, hjust = 0, data = tab_acc)

g_acc_cor <- ggplot(aes(x=accuracy_1, y=accuracy_2), data = NULL) +
  geom_abline(size = 0.3, colour = palette_primary["grey"]) +
  geom_point_nonwinners(data=tab_acc_nonwin, alpha=0.7, mapping=aes_col) +
  geom_point_winners(data=tab_acc_win, alpha=1, mapping=aes_col) +
  stat_aspear() +
  facet_grid(accuracy_nice_1 ~ accuracy_nice_2, switch = "both") +
  scale_x_acc() + scale_y_acc() + scale_colour_winners() +
  theme_cor

g_acc_cor
```


```{r,}
#------------------------------------------------------------------------------
# Part 2: Inter-X-metric correlations
#------------------------------------------------------------------------------

# Combinations of metrics to compare
x_levels <- c("X99", "X95", "X90", "X80")
x_types  <- fct_inorder(x_levels)
x_types_rev <- fct_inorder(rev(x_levels))
tab_x_base <- expand_grid(ranking = data_teams$prediction_track_ranking,
                            x_type_1 = x_types, x_type_2 = x_types_rev) %>%
  mutate(winning_rank = ifelse(ranking <= n_winners, ranking, "Other"))

# Add values
tab_teams_melt_x <- data_teams %>% 
  select(ranking = prediction_track_ranking, starts_with("X")) %>%
  gather(x_type, N, -ranking)
tab_x <- tab_x_base %>%
  inner_join(tab_teams_melt_x, 
             by = c("ranking" = "ranking", 
                    "x_type_1" = "x_type")) %>% 
  rename(N1 = N) %>%
  inner_join(tab_teams_melt_x, 
             by = c("ranking" = "ranking", 
                    "x_type_2" = "x_type")) %>% 
  rename(N2 = N) %>%
  mutate(x_type_1 = factor(x_type_1, levels = x_levels),
         x_type_2 = factor(x_type_2, levels = rev(x_levels)))
  
# Split into winners/nonwinners
tab_x_win <- tab_x %>% filter(winning_rank != "Other")
tab_x_nonwin <- tab_x %>% filter(winning_rank == "Other")

# Make plots
stat_xspear <- purrr::partial(stat_spear_no_p, label.x = n_labs*0.05, 
                              label.y = n_labs*0.95,
                              vjust = 1, hjust = 0, data = tab_x)

g_x_cor <- ggplot(aes(x=N1, y=N2), data = NULL) +
  geom_abline(size = 0.3, colour = palette_primary["grey"]) +
  geom_point_nonwinners(data=tab_x_nonwin, alpha=0.7, mapping=aes_col) +
  geom_point_winners(data=tab_x_win, alpha=1, mapping=aes_col) +
  stat_xspear() +
  facet_grid(x_type_1 ~ x_type_2, switch = "both") +
  scale_x_nlabs() + scale_y_nlabs() + scale_colour_winners() +
  theme_cor

g_x_cor
```

```{r,}
#------------------------------------------------------------------------------
# Part 3: X-vs-accuracy correlations (all teams)
#------------------------------------------------------------------------------

# Combinations to compare
tab_mix_base <- expand_grid(ranking = data_teams$prediction_track_ranking,
                            accuracy_type = tab_types$accuracy_type_base,
                            x_type = x_types_rev) %>%
  mutate(winning_rank = ifelse(ranking <= n_winners, ranking, "Other"))

# Add nice accuracy names
tab_mix_nice <- tab_mix_base %>%
  inner_join(tab_types,
             by = c("accuracy_type" = "accuracy_type_base")) %>% 
  rename(accuracy_nice = accuracy_type_nice_asc) %>%
  select(-accuracy_type_nice_desc)

# Add values
tab_mix <- tab_mix_nice %>%
  inner_join(tab_teams_melt, by = c("ranking", "accuracy_type")) %>%
  inner_join(tab_teams_melt_x, by = c("ranking", "x_type")) %>%
  mutate(x_type = factor(x_type, levels = x_levels))

# Split into winners/nonwinners
tab_mix_win <- tab_mix %>% filter(winning_rank != "Other")
tab_mix_nonwin <- tab_mix %>% filter(winning_rank == "Other")

# Make plots
stat_mspear <- purrr::partial(stat_spear_no_p, label.x = 0.05, 
                              label.y = n_labs*0.95,
                              vjust = 1, hjust = 0, data = tab_mix)
g_mix_cor_all <- ggplot(aes(x=accuracy, y=N), data = NULL) +
  geom_point_nonwinners(data=tab_mix_nonwin, alpha=0.7, mapping=aes_col) +
  geom_point_winners(data=tab_mix_win, alpha=1, mapping=aes_col) +
  stat_mspear() +
  facet_grid(x_type ~ accuracy_nice, switch = "both") +
  scale_x_acc() + scale_y_nlabs() + scale_colour_winners() +
  theme_cor
g_mix_cor_all
```

```{r,}
#------------------------------------------------------------------------------
# Part 4: X-vs-accuracy correlations (filtered)
#------------------------------------------------------------------------------

# Filter data to remove outliers
tab_mixf <- tab_mix %>% filter(N < n_labs)
tab_mixf_win <- tab_mixf %>% filter(winning_rank != "Other")
tab_mixf_nonwin <- tab_mixf %>% filter(winning_rank == "Other")

# Make plots
stat_fspear <- purrr::partial(stat_spear_no_p, label.x = 0.05, 
                              label.y = n_labs*0.95,
                              vjust = 1, hjust = 0, data = tab_mixf)
g_mix_cor_filtered <- ggplot(aes(x=accuracy, y=N), data = NULL) +
  geom_point_nonwinners(data=tab_mixf_nonwin, alpha=0.7, mapping=aes_col) +
  geom_point_winners(data=tab_mixf_win, alpha=1, mapping=aes_col) +
  stat_fspear() +
  facet_grid(x_type ~ accuracy_nice, switch = "both") +
  scale_x_acc() + scale_y_nlabs() + scale_colour_winners() +
  theme_cor

g_mix_cor_filtered
```

# Supplementary Figure: Sequences per lab

```{r,}
#==============================================================================
# SI FIGURES: SEQUENCES PER LAB
#==============================================================================

# Libraries
source("scripts/aux_format-plots.R")

# Input paths
data_path <- "data/lab-counts.csv"

# Output paths
out_path_all  <- "figures/si_sequences-per-lab.png"
out_path_test <- "figures/si_test-lab-composition.png"

# Output parameters
output_width_in <- fig_width_single_wide_in
output_asp <- 0.7

# Palettes
palette_test <- unname(palette_secondary[c("red", "blue")])

#------------------------------------------------------------------------------
# Import and process data
#------------------------------------------------------------------------------

# Import data
data <- import_csv(data_path)

# Pre-collapse meta-counts
data_all <- data %>%
  mutate(n_all = n_leaderboard + n_test + n_train) %>%
  group_by(n_seq = n_all) %>% summarise(n_labs = n())

# Test-set counts (post-collapse)
id_unkn <- max(data$lab_id_hidden, na.rm = TRUE) + 1
data_test <- data %>% filter(n_test > 0) %>%
  mutate(lab_id_hidden = ifelse(unknown_engineered, id_unkn, 
                                lab_id_hidden)) %>%
  group_by(lab_id_hidden, unknown_engineered) %>% 
  summarise(n_seq = sum(n_test), .groups = "drop") %>%
  mutate(lab_rank = row_number(desc(n_seq)),
         p_seq = n_seq/sum(n_seq),
         unk_eng = ifelse(unknown_engineered,
                          "Unknown Engineered", "Other lab"),
         unk_eng = factor(unk_eng, 
                          levels = c("Unknown Engineered", "Other lab"))) %>%
  select(lab_rank, p_seq, unk_eng)
```

```{r,}
#------------------------------------------------------------------------------
# Plot pre-collapse meta-counts
#------------------------------------------------------------------------------

g_ccount <- ggplot(data_all, aes(x=n_seq, y=n_labs)) +
  geom_line(size = 0.5, colour = palette_primary_dark["grey"]) + 
  geom_point(shape = 16, size = 2, colour = palette_primary_dark["grey"]) +
  scale_x_log10(name="Plasmid entries per lab") +
  scale_y_continuous(name="Number of labs", minor_breaks = NULL) +
  theme_base + theme(aspect.ratio = 1/2)
g_ccount
```

```{r,}
#------------------------------------------------------------------------------
# Plot test-set composition
#------------------------------------------------------------------------------

g_test <- ggplot(data_test, aes(x=lab_rank, y=p_seq, colour=unk_eng)) +
  geom_line(size = 0.2, colour = palette_test[2]) +
  geom_point() +
  scale_x_log10("Lab ranking (by number of sequences in test set)") +
  scale_y_continuous(name = "Percentage of test set",
                     labels = function(x) as.integer(x*100),
                     limits = c(0,0.08)) +
  scale_colour_manual(name = "Lab",
                      values = palette_test) +
  theme_base + theme(aspect.ratio = 1/2)
g_test
```

# Supplementary Figure: Ensemble Top-N Accuracy

#==============================================================================
# SI FIGURE: ENSEMBLE TOP-N ACCURACY
#==============================================================================

```{r,}
# Libraries & formatting
source("scripts/aux_format-plots.R")

# Input paths
path_acc_teams <- "data/team-top-n-accuracy.csv"
path_acc_ens   <- "data/key-model-top-n-accuracy.csv"

# Output paths
out_path <- "figures/si_ensemble-top-n.png"

# Output parameters
output_width_in <- fig_width_single_wide_in
output_asp <- 0.7

# Define palettes, scales, themes
palette_model <- unname(palette_primary[c("yellow", "green1")])
scale_colour_model <- purrr::partial(scale_colour_manual,
                                     values = palette_model,
                                     name = "Model:")
theme_wide <- theme_base + theme(aspect.ratio = 1/2)

#------------------------------------------------------------------------------
# Import and process data
#------------------------------------------------------------------------------

# Import data
data_acc_ens   <- import_csv(path_acc_ens) %>%
  filter(grepl("Ensemble", source)) %>% select(-source)
data_acc_teams <- import_csv_matrix(path_acc_teams)

# Extract 1st-place accuracies
data_acc_geac_1st <- data_acc_teams[which.max(data_acc_teams[,10]),]

# Combine accuracies
data_acc_n <- data_acc_ens %>% rename(`GEAC Ensemble` = accuracy_top_n) %>%
  mutate(`GEAC (1st place)` = data_acc_geac_1st) %>%
  gather(model, accuracy_top_n, -n) %>%
  mutate(model = fct_inorder(model))
```

```{r,}
#------------------------------------------------------------------------------
# Make plot
#------------------------------------------------------------------------------

g_n <- ggplot(data_acc_n, aes(x=n, y=accuracy_top_n, colour = model)) +
  geom_line_thin() + geom_point_fixed() +
  scale_x_log10(name = "N") +
  scale_y_cont_nominor(name = "Top-N Accuracy (%)", limits = c(0.8, 1),
                       breaks = seq(0,1,0.05),
                       labels = function(y) y * 100) +
  scale_colour_model() + theme_wide
g_n
```



# Supplementary Figure: Key models

#==============================================================================
# KEY MODEL SUPPLEMENTARY FIGURES
#==============================================================================
```{r,}
# Libraries & formatting
source("scripts/aux_format-plots.R")

# Parameters
source_names_raw <- c("BLAST (new)", "BLAST (Alley et al., original)",
                  "BLAST (Alley et al., corrected)", "Nielsen & Voigt (2018)", "Alley et al. (2020)",
                  "GEAC (1st place)", "GEAC (2nd place)", "GEAC (3rd place)",
                  "GEAC (4th place)", "GEAC Ensemble")
source_names_plot <- c(NA, "BLAST (original)",
                      "BLAST (modified)", "Nielsen & Voigt (2018)",
                      "Alley et al. (2020)",
                      "GEAC (1st place)", "GEAC (2nd place)", "GEAC (3rd place)",
                      "GEAC (4th place)", "GEAC Ensemble")
levels_unkn <- c("All labs", "Known labs only", "Unknown Engineered only")
unk_eng_true_frequency <- 0.07541186

# Input paths
path_key <- "data/key-model-stats.csv"
path_unkn <- "data/key-model-unknown-engineered-stats.csv"

# Output parameters
out_path_acc     <- "figures/si_key-model-accuracy.png"
out_path_x       <- "figures/si_key-model-xmetrics.png"
out_path_cal     <- "figures/si_key-model-calibration.png"
out_path_unkn    <- "figures/si_key-model-unknown-engineered.png"
out_path_f1      <- "figures/si_key-model-f1.png"
out_width_full   <- page_width_in
out_width_single <- fig_width_single_wide_in
out_asp_acc      <- 1.5
out_asp_x        <- 1.5
out_asp_cal      <- 0.82
out_asp_unkn     <- 0.9
out_asp_f1       <- 0.82

# Themes, palettes, geoms
palette_calibration <- unname(c(palette_primary["grey"],
                                palette_primary_dark["grey"]))
palette_f1 <- unname(c(palette_primary_mid["grey"],
                       palette_primary_dark["grey"],
                       "black"))
palette_nlabs <- unname(palette_primary_darkmid["grey"])
scale_fill_cal <- purrr::partial(scale_fill_manual, name = "Metric:",
                                 values = palette_calibration)
scale_fill_f1 <- purrr::partial(scale_fill_manual, name = "Metric:",
                                values = palette_f1)
geom_hline_nlabs <- purrr::partial(geom_hline, yintercept = n_labs,
                                   colour = palette_nlabs, size = 0.3,
                                   linetype = "dashed")
annot_nlabs <- purrr::partial(annotate, geom = "text", x = Inf,
                              label = "All labs", hjust = 1, vjust = 1,
                              size = rescale_font(fontsize_base),
                              colour = palette_nlabs)
annot_nlabs_lin <- purrr::partial(annot_nlabs, y = n_labs-40)
annot_nlabs_log <- purrr::partial(annot_nlabs, y = n_labs*0.87)

#------------------------------------------------------------------------------
# Import data
#------------------------------------------------------------------------------

data_key <- import_csv(path_key)
data_unkn <- import_csv(path_unkn)

#------------------------------------------------------------------------------
# Process data
#------------------------------------------------------------------------------

# Modify sources and source names for display
tab_sources <- tibble(source = source_names_raw,
                      source_new = source_names_plot)
rename_sources <- function(tab){
  inner_join(tab, tab_sources, by = "source") %>%
    filter(!is.na(source_new)) %>% select(-source) %>%
    select(source = source_new, everything()) %>%
    mutate(source = factor(source, levels = source_names_plot[-1])) %>%
    arrange(source)
}

data_key_named <- rename_sources(data_key)
data_unkn_named <- rename_sources(data_unkn) %>%
  mutate(data_subset = factor(data_subset, levels = levels_unkn))

#------------------------------------------------------------------------------
# Accuracy
#------------------------------------------------------------------------------

scale_x_key <- purrr::partial(scale_x_discrete, name = "Model")
theme_col_bare_wide <- theme_col + 
  theme(aspect.ratio = 1/2,
        axis.title.x.bottom = element_blank())

tab_accuracy <- data_key_named %>% select(source, starts_with("accuracy")) %>%
  gather(metric, accuracy, -source) %>%
  mutate(metric = sub("accuracy_top", "Top ", metric),
         metric = fct_inorder(metric))
aes_col <- purrr::partial(aes, x=source, fill=metric)
```


```{r,}
g_accuracy <- ggplot(tab_accuracy, aes_col(y=accuracy)) +
  geom_col_mid(position = "dodge") +
  scale_x_key() + scale_y_percent(name = "Accuracy (%)") +
  scale_fill_accuracy() + theme_col_bare_wide
g_accuracy

g_error <- ggplot(tab_accuracy, aes_col(y=1-accuracy)) +
  geom_col_mid(position = "dodge") +
  scale_x_key() + theme_col_bare_wide +
  scale_y_cont_nominor(name = "Misclassification rate (%)", expand = c(0,0),
                       labels = function(y) as.integer(y * 100),
                       limits = c(0,0.5), breaks = seq(0,1,0.1)) +
  scale_fill_accuracy() + theme_col_bare_wide
g_error

g_acc_out <- wrap_plots(g_accuracy, g_error, ncol = 1) +
  plot_annotation(tag_levels = "a") + plot_layout(guides = "collect") &
  theme_legend
g_acc_out
```

```{r,}
#------------------------------------------------------------------------------
# X-metrics
#------------------------------------------------------------------------------

tab_x <- data_key_named %>% select(source, starts_with("X")) %>%
  gather(metric, N, -source) %>% arrange(desc(metric)) %>%
  mutate(metric = fct_inorder(metric))

plot_x <- function(data){
  ggplot(data, aes_col(y=N)) + geom_col_mid(position = "dodge") +
    geom_hline_nlabs() + scale_x_key() + scale_fill_x() + theme_col_bare_wide
}

g_x_all <- plot_x(tab_x) + scale_y_log_nominor(expand=c(0,0)) +
  annot_nlabs_log()
g_x_all

g_x_large <- plot_x(filter(tab_x, metric %in% c("X99", "X95"))) +
  scale_y_cont_nominor(expand=c(0,0), breaks = seq(0,2000,400)) +
  annot_nlabs_lin()
g_x_large

g_x_out <- wrap_plots(g_x_all, g_x_large + guides(fill = FALSE), ncol = 1) +
  plot_annotation(tag_levels = "a") + plot_layout(guides = "collect") &
  theme_legend
g_x_out
```

```{r,}
#------------------------------------------------------------------------------
# Calibration
#------------------------------------------------------------------------------

tab_cal <- data_key_named %>% select(source, contains("calibration")) %>%
  gather(metric, error, -source) %>%
  mutate(metric = toupper(sub(".._calibration_error", "ce", metric)),
         metric = fct_inorder(metric))
  
g_cal <- ggplot(tab_cal, aes_col(y=error)) +
  geom_col_mid(position = "dodge") + scale_x_key() +
  scale_y_cont_nominor(name = "Calibration error (%)", expand = c(0,0),
                       labels = function(y) as.integer(y * 100),
                       limits = c(0,0.4), breaks = seq(0,1,0.1)) +
  scale_fill_cal() + theme_col_bare_wide
g_cal

g_cal_out <- wrap_plots(g_cal, ncol = 1) +
  plot_layout(guides = "collect") & theme_legend
g_cal_out
```

```{r,}
#------------------------------------------------------------------------------
# Precision/recall/F1
#------------------------------------------------------------------------------

tab_f1 <- data_key_named %>% select(source, starts_with("macro")) %>%
  rename(Precision = macro_precision, Recall = macro_recall,
         `F1 score` = macro_f1) %>%
  gather(metric, value, -source) %>%
  mutate(metric = fct_inorder(metric))

g_f1 <- ggplot(tab_f1, aes_col(y=value)) +
  geom_col_mid(position = "dodge") + scale_x_key() +
  scale_y_proportion(name = NULL) + scale_fill_f1() +
  theme_col_bare_wide
g_f1

g_f1_out <- wrap_plots(g_f1, ncol = 1) +
  plot_layout(guides = "collect") & theme_legend
g_f1_out
```

# Supplementary Figure: Unknown engineered
```{r,}
#------------------------------------------------------------------------------
# Unknown Engineered
#------------------------------------------------------------------------------

aes_unkn <- purrr::partial(aes, x = source, fill = data_subset)

# Top-10 accuracy
g_unkn_acc_top10 <- ggplot(data_unkn_named, aes_unkn(y=accuracy_top10)) +
  geom_col_mid(position = "dodge") + scale_x_key() +
  scale_y_percent(name = "Top-10 accuracy (%)") + 
  scale_fill_unkn() + theme_col_bare_wide
g_unkn_acc_top10

# Top-1 accuracy
g_unkn_acc_top1 <- ggplot(data_unkn_named, aes_unkn(y=accuracy_top1)) +
  geom_col_mid(position = "dodge") + scale_x_key() +
  scale_y_percent(name = "Top-1 accuracy (%)") + 
  scale_fill_unkn() + theme_col_bare_wide
g_unkn_acc_top1

# Rate of Unknown Engineered in top-10
g_unkn_in_top10 <- ggplot(data_unkn_named, aes_unkn(y=unk_eng_in_top10)) +
  geom_col_mid(position = "dodge") + scale_x_key() +
  geom_hline(yintercept = unk_eng_true_frequency, colour = palette_secondary["red"],
             linetype = "dashed", size = 0.3) +
  scale_y_percent(name = "Mean frequency of\nUnk. Eng. in top 10 guesses") + 
  scale_fill_unkn() + theme_col_bare_wide
g_unkn_in_top10

# Average rank of unknown engineered
g_unkn_rank <- ggplot(data_unkn_named, aes_unkn(y=unk_eng_avg_rank)) +
  geom_col_mid(position = "dodge") + scale_x_key() +
  scale_y_cont_nominor(name = "Geometric mean of\nUnk. Eng. ranking",
                       expand = c(0,0)) + 
  scale_fill_unkn() + theme_col_bare_wide
g_unkn_rank

g_unkn_out <- wrap_plots(g_unkn_in_top10, g_unkn_rank, g_unkn_acc_top10,
                      g_unkn_acc_top1, ncol = 2) +
  plot_annotation(tag_levels = "a") + plot_layout(guides = "collect") &
  theme_legend
g_unkn_out
```

# Supplementary Figure: X99 vs X95
#==============================================================================
# SUPPLEMENTARY FIGURE: X99 VS X95
#==============================================================================

```{r,}

# Libraries & formatting
source("scripts/aux_format-plots.R")

# Input paths (team data)
path_teams <- "data/team-stats.csv"

# Output paths
out_path_xrel <- "figures/si_x99-vs-x95.png"

# Output parameters
output_width_xrel <- fig_width_single_narrow_in
output_asp_xrel   <- 0.9

#------------------------------------------------------------------------------
# Define palettes and scales
#------------------------------------------------------------------------------

# Winners vs nonwinners
palette_win <- unname(palette_primary_dark[c("grey", "green2")])
palette_win_dec <- c(palette_win[2], rep("black", 10))
scale_fill_win <- purrr::partial(scale_fill_manual,
                               name = "Metric:",
                               values = palette_win)

#------------------------------------------------------------------------------
# Read data
#------------------------------------------------------------------------------

data_teams <- suppressMessages(read_csv(path_teams))

#------------------------------------------------------------------------------
# Add decile labels
#------------------------------------------------------------------------------

tab_teams_labelled <- data_teams %>% group_by(ranking_decile) %>%
  mutate(decile_rank_min = min(prediction_track_ranking),
         decile_rank_max = max(prediction_track_ranking),
         decile_label = paste(decile_rank_min, 
                              decile_rank_max, sep=" to "),
         decile_label_long = paste("Rank", decile_label),
         decile_label = fct_inorder(decile_label),
         decile_label_long = fct_inorder(decile_label_long))

#------------------------------------------------------------------------------
# X99 vs X95
#------------------------------------------------------------------------------

# Compute ratio
tab_teams_xrel <- mutate(tab_teams_labelled,
                         xrel = X99/X95)

# Add separate winners column
tab_teams_xrel_win <- tab_teams_xrel %>% 
  filter(prediction_track_ranking <= n_winners) %>%
  mutate(ranking_decile = 0, 
         decile_label = paste(1, n_winners, sep = " to "),
         decile_label_long = paste("Rank", decile_label)) %>%
  bind_rows(tab_teams_xrel) %>%
  mutate(decile_label = fct_inorder(decile_label),
         decile_label_long = fct_inorder(decile_label_long),
         winner = (ranking_decile == 0))

# Summarise by decile
tab_teams_xrel_summ <- tab_teams_xrel_win %>%
  group_by(ranking_decile) %>%
  summarise(geomean_xrel = geomean(xrel),
            ranking_decile = ranking_decile[1],
            decile_label = decile_label[1],
            decile_label_long = decile_label_long[1],
            winner = winner[1])
```


```{r,}
g_xrel_deciles <- ggplot(tab_teams_xrel_summ,
                         aes(x=decile_label, y = geomean_xrel,
                             fill = winner)) + 
  scale_x_discrete(name = "Innovation Track ranking") +
  scale_y_cont_nominor(name = "X99  X95 (geometric mean)", limits = c(0,10),
                       breaks = seq(0,100,2), expand = c(0,0)) +
  scale_fill_win(guide = FALSE) +
  geom_col() + theme_col + theme(
    axis.text.x.bottom = element_text(colour = palette_win_dec),
  )
g_xrel_deciles
```

# Session Information

```{r,}
sessionInfo()
```

